# Chapter 3: Data Visualization

```{r}
library(tidyverse)
```

## 3.2 First Steps

1\. Run ggplot(data = mpg). What do you see?

```{r}
ggplot(data = mpg)
```

Nothing

2\. How many rows are in mtcars? How many columns?

```{r}
mpg
```

234 x 11

3\. What does the drv variable describe? Read the help for ?mpg to find out.

the type of drive train, where f = front-wheel drive, r = rear wheel drive, 4 = 4wd

4\. Make a scatterplot of hwy versus cyl.

```{r}
ggplot(mpg) +
  geom_point(aes(x = hwy, y = cyl))
```

5\. What happens if you make a scatterplot of class versus drv? Why is the plot not useful?

```{r}
ggplot(mpg) +
  geom_point(aes(x = class, y = drv))
```

Both discrete

## 3.3 Aesthetics

1\. What's gone wrong with this code? Why are the points not blue?

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = "blue"))
```

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy), color = "blue")
```

2\. Which variables in mpg are categorical? Which variables are continuous? (Hint: type ?mpg to read the documentation for the data set) How can you see this information when you run mpg?

```{r}
mpg
```

displ, year, cty, hwy

3\. Map a continuous variable to color, size, and shape. How do these aesthetics behave differently for categorical versus continuous variables?

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = year))
```

Bar color happens

4\. What happens if you map the same variable to multiple aesthetics?

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = hwy))
```

Repeated variable happens

5\. What does the stroke aesthetic do? What shapes does it work with? (Hint: use ?geom_point.)

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy), stroke = 10)
```

Makes points bigger

6\. What happens if you map an aesthetic to something other than a variable name, like aes(color = displ \< 5)?

```{r}

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color= displ < 5))
```

True and False

## 3.4 Facet

1\. What happens if you facet on a continuous variable?

```{r}
mpg
```

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = cty)) + 
  facet_wrap(~displ, nrow = 4)
```

Creates for each value

2\. What do the empty cells in a plot with facet_grid(drv \~ cyl) mean? How do they relate to this plot?

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = drv, y = cyl)) +
  facet_grid(drv ~ cyl)
```

No values, correspond with the original.

3\. What plots does the following code make? What does . do?

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_grid(drv ~ .)
```

Removed grid

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_grid(. ~ cyl)
```

Removed rows

4\. Take the first faceted plot in this section:

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_wrap(~ class, nrow = 2)
```

What are the advantages to using faceting instead of the color aesthetic? What are the disadvantages? How might the balance change if you had a larger dataset?

Harder to compare, but better when data set gets larger.

5\. Read ?facet_wrap. What does nrow do? What does ncol do? What other options control the layout of the individual panels? Why doesn't facet_grid() have nrow and ncol variables?

```{r}
facet_wrap
```

Controls how many columns, facet_grid() doesn't because ncol depends on a variable

6\. When using facet_grid() you should usually put the variable with more unique levels in the columns. Why?

More Space

## 3.5 Geometric Objects

1\. What geom would you use to draw a line chart? A boxplot? A histogram? An area chart?

```{r}
ggplot(mpg) + 
  geom_boxplot(aes(x = class, y = displ))

ggplot(mpg) +
  geom_smooth(aes(x = displ, y = hwy))

ggplot(mpg) + 
  geom_histogram(aes(x = displ))
```

2\. Run this code in your head and predict what the output will look like. Then, run the code in R and check your predictions:

```{r}
ggplot(
  data = mpg,
  mapping = aes(x = displ, y = hwy, color = drv)) +
  geom_point() +
  geom_smooth(se = FALSE)
```

Points and line without se

3\. What does show.legend = FALSE do? What happens if you remove it? Why do you think I used it earlier in the chapter?

```{r}
ggplot(
  data = mpg,
  mapping = aes(x = displ, y = hwy, color = drv)) +
  geom_smooth(se = FALSE, , show.legend = FALSE)
```

4\. What does the se argument to geom_smooth() do?

It gets rid of the shading area.

5\. Will these two graphs look different? Why/why not?

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point() + 
  geom_smooth()

ggplot() + 
  geom_point(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_smooth(data = mpg, mapping = aes(x = displ, y = hwy))
```

They will be the same

6\. Re-create the R code necessary to generate the following graphs.

```{r}
ggplot(mpg, aes(x = displ, y = hwy)) +
  geom_point(stroke = 3) +
  geom_smooth(se = FALSE)
```

```{r}
ggplot(mpg, aes(x = displ, y = hwy)) +
  geom_point(stroke = 3) +
  geom_smooth(aes(group = drv), se = FALSE)
```

```{r}
ggplot(mpg, aes(x = displ, y = hwy, color = drv)) +
  geom_point(stroke = 3) +
  geom_smooth(se = FALSE)
```

```{r}
ggplot(mpg, aes(x = displ, y = hwy)) +
  geom_point(aes(color = drv), stroke = 3) +
  geom_smooth(se = FALSE)
```

```{r}
ggplot(mpg, aes(x = displ, y = hwy)) +
  geom_point(aes(color = drv), stroke = 3) +
  geom_smooth(aes(linetype = drv), se = FALSE)
```

```{r}
ggplot(mpg, aes(x = displ, y = hwy)) +
  geom_point(aes(color = drv), colour = "white", stroke = 3) + 
  geom_point(aes(color = drv), stroke = 2)
```

## 3.6 Statistical Transformations

1\. What is the default geom associated with stat_summary()? How could you rewrite the previous plot to use that geom function instead of the stat function?

```{r}
ggplot(diamonds) +
  stat_summary(aes(x = cut, y = price))
```

Weird version of boxplot - pointrange

2\. What does geom_col() do? How is it different to geom_bar()?

```{r}
ggplot(diamonds) +
  geom_col(aes(x = cut, y = price))
```

Uses two variables as opposed to one, stat function becomes sum

3\. Most geoms and stats come in pairs that are almost always used in concert. Read through the documentation and make a list of all the pairs. What do they have in common?

| **geom**                | **stat**                |
|-------------------------|-------------------------|
| `geom_bar()`            | `stat_count()`          |
| `geom_bin2d()`          | `stat_bin_2d()`         |
| `geom_boxplot()`        | `stat_boxplot()`        |
| `geom_contour_filled()` | `stat_contour_filled()` |
| `geom_contour()`        | `stat_contour()`        |
| `geom_count()`          | `stat_sum()`            |
| `geom_density_2d()`     | `stat_density_2d()`     |
| `geom_density()`        | `stat_density()`        |
| `geom_dotplot()`        | `stat_bindot()`         |
| `geom_function()`       | `stat_function()`       |
| `geom_sf()`             | `stat_sf()`             |
| `geom_sf()`             | `stat_sf()`             |
| `geom_smooth()`         | `stat_smooth()`         |
| `geom_violin()`         | `stat_ydensity()`       |
| `geom_hex()`            | `stat_bin_hex()`        |
| `geom_qq_line()`        | `stat_qq_line()`        |
| `geom_qq()`             | `stat_qq()`             |
| `geom_quantile()`       | `stat_quantile()`       |

Similar names.

4\. What variables does stat_smooth() compute? What parameters control its behavior?

```{r}
ggplot(diamonds) + 
  stat_smooth(aes(x = price, y = carat))
```

-   `after_stat(y)` *or* `after_stat(x)`\
    Predicted value.

-   `after_stat(ymin)` *or* `after_stat(xmin)`\
    Lower pointwise confidence interval around the mean.

-   `after_stat(ymax)` *or* `after_stat(xmax)`\
    Upper pointwise confidence interval around the mean.

-   `after_stat(se)`\
    Standard error.

5\. In our proportion bar chart, we need to set group = 1. Why? In other words what is the problem with these two graphs?

```{r}
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, y = after_stat(prop)))
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, fill = color, y = after_stat(prop)))
```

```{r}
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, y = after_stat(prop), group = 1))
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, fill = color, y = ..count.. / sum(..count..)))
```

## 3.7 Position Adjustments

1\. What is the problem with this plot? How could you improve it?

```{r}
ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) + 
  geom_point()
```

```{r}
ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) + 
  geom_point(position = "jitter")
```

2\. What parameters to geom_jitter() control the amount of jittering?

width and height

3\. Compare and contrast geom_jitter() with geom_count().

```{r}
ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) + 
  geom_jitter()
```

```{r}
ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) + 
  geom_count()
```

geom_count counts the density.

4\. What's the default position adjustment for geom_boxplot()? Create a visualization of the mpg dataset that demonstrates it.

```{r}
ggplot(mpg) + 
  geom_boxplot(aes(x = cyl, y = displ, fill = fl))
```

## 3.8 Coordinates

1\. Turn a stacked bar chart into a pie chart using coord_polar().

```{r}
ggplot(diamonds) +
  geom_bar(aes(x = cut, fill = color), position = "fill") + 
  
  coord_polar(theta = "y")
```

lol

2\. What does labs() do? Read the documentation.

Modifies axis, legend and plot labels

3\. What's the difference between coord_quickmap() and coord_map()?

Quickmap is a quick approximation that preserves straight lines.

4\. What does the following plot tell you about the relationship between city and highway mpg? Why is coord_fixed() important? What does geom_abline() do?

```{r}
ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) +
  geom_point() + 
  geom_abline() +
  coord_fixed()
```

coord_fix so you can compare, abline is a reference line, currently at y = x, the relationship is y = x +c
